
Load required libraries

```{r eval=FALSE}

library(Seurat)
library(tidyverse)
library(cowplot)
library(patchwork)
library(WGCNA)
library(Matrix)
library(viridis)
library(harmony)
library(RColorBrewer)
library(ggpubr)
library(tictoc)
library(RColorBrewer)
library(Hmisc)
library(corrplot)
library(enrichR)
library(GeneOverlap)
library(grid)
library(gridExtra)
library(igraph)
library(ggrepel)
library(hdWGCNA)
enableWGCNAThreads(nThreads = 8)
theme_set(theme_cowplot())
set.seed(12345)

# Custom ST plotting functions
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

# output directories
data_dir <- "data/"
fig_dir <- 'figures/'

```

Run network analysis iteratively for each major cell type

```{r eval=FALSE}

# load the processed Seurat object for the Zhou et al dataset
# with only control brain samples
seurat_obj <- readRDS(file='/dfs7/swaruplab/smorabit/analysis/scWGCNA/data/Zhou_control_scWGCNA.rds')

# get list of each cell type
groups <- as.character(unique(seurat_obj$cell_type))

for(i in 1:length(groups)){

  print(i)
  cur_group <- as.character(groups[i])

  # only run metacell aggregation one time
  if(i == 1){
    print('Setting up and running metacells')
    # setup training data for scWGCNA:
    seurat_obj <- SetupForWGCNA(
      seurat_obj,
      gene_select = "fraction",
      fraction = 0.05,
      wgcna_name = cur_group
    )

    # construct metacells:
    seurat_obj <- MetacellsByGroups(
      seurat_obj = seurat_obj,
      group.by = c("cell_type", "Sample"),
      k = 25,
      max_shared = 12,
      reduction = 'harmony',
      ident.group = 'cell_type'
    )

    seurat_obj <- NormalizeMetacells(seurat_obj)

  } else{

    seurat_obj  <- SetupForWGCNA(
      seurat_obj,
      gene_select = "fraction",
      fraction = 0.05,
      wgcna_name = cur_group,
      metacell_location = groups[1]
    )
  }

  # setup expression matrix
  seurat_obj <- SetDatExpr(
    seurat_obj,
    group.by='cell_type',
    group_name = cur_group
  )

  # construct network
  seurat_obj <- TestSoftPowers(seurat_obj)
  seurat_obj <- ConstructNetwork(seurat_obj, tom_name = cur_group)
  seurat_obj <- ModuleEigengenes(seurat_obj, group.by.vars="Sample", wgcna_name = cur_group)
  seurat_obj <- ModuleConnectivity(seurat_obj, group.by='cell_type', group_name = cur_group, wgcna_name = cur_group)

  # rename the modules
  seurat_obj <- ResetModuleNames(
    seurat_obj,
    new_name = paste0(cur_group, '-M'),
    wgcna_name=cur_group
  )

}

# save the output
saveRDS(seurat_obj, file=paste0(data_dir, 'Zhou_scWGCNA_all_celltypes.rds'))

```


Write the module assignments to file

```{r eval=FALSE}

groups <- as.character(unique(seurat_obj$cell_type))
modules <- do.call(rbind, lapply(groups, function(x){
  m <- GetModules(seurat_obj, wgcna_name=x) %>% subset(module != 'grey')
  m$group <- x
  m %>% dplyr::select(c(gene_name, module, color, group))
}))

write.csv(modules, row.names=FALSE, quote=FALSE, file=paste0(data_dir, 'brain_modules.csv'))

```


Run enrichR pathway enrichment in each cell type

```{r eval=FALSE}

library(enrichR)

dbs <-c('GO_Biological_Process_2021','GO_Cellular_Component_2021','GO_Molecular_Function_2021','LINCS_L1000_Chem_Pert_up','LINCS_L1000_Chem_Pert_down', 'WikiPathway_2021_Human', 'KEGG_2021_Human')

# compute GO terms:
groups <- as.character(unique(seurat_obj$cell_type))
enrich_list <- list()
for(cur_group in groups){
  seurat_obj <- RunEnrichr(seurat_obj, dbs=dbs, wgcna_name=cur_group)
  enrichr_df <- GetEnrichrTable(seurat_obj, wgcna_name=cur_group) %>% subset(P.value < 0.05)
  enrichr_df$cell_type <- cur_group
  enrich_list[[cur_group]] <- enrichr_df
}

enrich_df <- do.call(rbind, enrich_list)
write.table(enrichr_df, quote=FALSE, sep='\t', row.names=FALSE, file=paste0(data_dir, 'brain_enrichr.tsv'))


```

Network visualizations and module eigengene plots


```{r eval=FALSE}

library(reshape2)
library(igraph)

groups <- as.character(unique(seurat_obj$cell_type))

# plot the dendrograms
pdf(paste0(fig_dir, "dendro.pdf"),height=3, width=6)
for(cur_group in groups){
    PlotDendrogram(seurat_obj, main=paste0(cur_group, ' Dendrogram'), wgcna_name=cur_group)
}
dev.off()

# module network plot
for(cur_group in groups){
  ModuleNetworkPlot(
    seurat_obj,
    mods = "all",
    outdir = paste0(fig_dir, cur_group, '_hubNetworks/'),
    wgcna_name = cur_group
  )
}

# module umap plot
plot_list <- list()
for(cur_group in groups){

  print(cur_group)

  # compute the module UMAPs
  seurat_obj <- RunModuleUMAP(
    seurat_obj,
    n_hubs = 5,
    n_neighbors=10,
    min_dist=0.3,
    spread=2,
    wgcna_name = cur_group,
    target_weight=0.1,
    supervised=TRUE
  )

  # get the hub gene UMAP table from the seurat object
  umap_df <- GetModuleUMAP(seurat_obj, wgcna_name=cur_group)

  # plot with ggplot
  plot_df <- umap_df

  # compute coordinates for cluster labels
  centroid_df <- data.frame()
  for(cur_cluster in unique(plot_df[['module']])){
    cur_meta <- plot_df[plot_df[['module']] == cur_cluster,]
    df <- data.frame(
      cluster = cur_cluster,
      UMAP1 = mean(cur_meta$UMAP1),
      UMAP2 = mean(cur_meta$UMAP2)
    )
    centroid_df <- rbind(centroid_df, df)
  }

  # plot with ggplot
  p <- ggplot(umap_df, aes(x=UMAP1, y=UMAP2)) +
    ggrastr::rasterise(geom_point(
     color=umap_df$color,
     size=umap_df$kME*2
   ), dpi=500, scale=0.5) +
    umap_theme() +
    theme(
      plot.margin = margin(0,0,0,0),
      plot.title = element_text(hjust=0.5)
    ) + ggtitle(cur_group) +
    ggrepel::geom_text_repel(data = centroid_df, label=centroid_df$cluster, color='black', max.overlaps=Inf, size=3)

  plot_list[[cur_group]] <- p


  pdf(paste0(fig_dir, cur_group, '_hubgene_umap_ggplot.pdf'), width=5, height=5)
  print(p)
  dev.off()

  hub_genes <- GetHubGenes(seurat_obj, 3, wgcna_name=cur_group)

  # add annotation
  anno_genes <- hub_genes$gene_name
  plot_df$anno <- ifelse(plot_df$gene %in% anno_genes, umap_df$gene, '')

  plot_df_anno <- subset(plot_df, anno != '')
  p <-  plot_df %>%
    ggplot(aes(x=UMAP1, y=UMAP2, color=module)) +
    ggrastr::rasterise(
      geom_point(
        inherit.aes=FALSE,
        data=plot_df,
        aes(x=UMAP1, y=UMAP2, color=module),
        color=plot_df$color,
        size=plot_df$kME*2,
      ), dpi=500, dpi_scale=0.5) +
    geom_point(
      inherit.aes = FALSE,
      data = plot_df_anno,
      shape=21, color='black',
      fill=plot_df_anno$color,
      size=plot_df_anno$kME*2,
      aes(x=UMAP1, y=UMAP2, fill=module)
    ) +
    # add labels
    ggrepel::geom_text_repel(data = centroid_df, label=centroid_df$cluster, color='black', max.overlaps=Inf, size=3, fontface='bold') +

    geom_text_repel(label=plot_df$anno, max.overlaps=Inf, color='black', fontface='italic', size=2) +
    umap_theme() + NoLegend() +
    coord_equal() +
    theme(
      plot.margin = margin(0,0,0,0)
    )

  pdf(paste0(fig_dir, cur_group, '_hubgene_umap_ggplot.pdf'), width=6, height=6)
  print(p)
  dev.off()

  # plot with igraph
  pdf(paste0(fig_dir, cur_group,'_hubgene_umap_igraph.pdf'), width=7, height=7)
  ModuleUMAPPlot(
    seurat_obj,
    edge.alpha=0.5,
    sample_edges=TRUE,
    keep_grey_edges=FALSE,
    edge_prop=0.075, # taking the top 20% strongest edges in each module
    #label_genes = label_genes,
    label_hubs=3, # how many hub genes to plot per module?
    wgcna_name = cur_group
  )
  dev.off()

}

pdf(paste0(fig_dir, 'hubgene_umap_ggplot_all.pdf'), width=9, height=6)
wrap_plots(plot_list, ncol=3)
dev.off()


# plot the featureplots
p_ex <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=100, alpha=1, raster_scale=0.25, wgcna_name='EX', restrict_range=FALSE)

p_inh <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=100, alpha=1,raster_scale=0.25, wgcna_name='INH', restrict_range=FALSE)

p_odc <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=100, alpha=1, raster_scale=0.25, wgcna_name='ODC', restrict_range=FALSE)

p_opc <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=100, alpha=1, raster_scale=0.25, wgcna_name='OPC', restrict_range=FALSE)

p_asc <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=100, alpha=1, raster_scale=0.25, wgcna_name='ASC', restrict_range=FALSE)

p_mg <- ModuleFeaturePlot(seurat_obj, order='shuffle', raster=TRUE, raster_dpi=200, alpha=1, raster_scale=0.25, wgcna_name='MG', restrict_range=FALSE)

plot_list <- c(
  p_ex, p_inh, p_odc, p_opc, p_asc, p_mg
)


pdf(paste0(fig_dir, "ME_featureplots.pdf"),height=20, width=20)
wrap_plots(plot_list, ncol=11)
dev.off()

```


Module Preservation analysis

```{r eval=FALSE}

# load Morabito & Miyoshi et al 2021 dataset
seurat_query <- readRDS(file='/dfs7/swaruplab/smorabit/analysis/scWGCNA/data/Swarup_2021.rds')

seurat_query <- ProjectModules(
  seurat_obj = seurat_query,
  seurat_ref = seurat_obj,
  wgcna_name = "ODC",
  wgcna_name_proj="ODC_projected"
)

plot_list <- ModuleFeaturePlot(seurat_query, order=TRUE, raster=TRUE, raster_dpi=400, alpha=1, restrict_range=FALSE, raster_scale=0.25)
pdf("figures/ODC_featureplot_MEs_morabito_2021.pdf",height=10, width=4)
wrap_plots(plot_list, ncol=2)
dev.off()

seurat_query <- SetDatExpr(
  seurat_query,
  group_name = "ODC",
  group.by = "cell_type",
  use_metacells = FALSE,
  assay = 'RNA'
)

seurat_query <- ModulePreservation(
  seurat_query,
  seurat_ref = seurat_obj,
  name="ODC_projected",
  wgcna_name = 'ODC_projected',
  wgcna_name_ref = 'ODC',
  verbose=3,
  n_permutations=100
)

library(ggrepel)
plot_list <- PlotModulePreservation(
  seurat_query,
  name="ODC_projected",
  statistics = "summary",
  label=FALSE
)

p <- plot_list[[2]] + theme(
  axis.line.x = element_blank(),
  axis.line.y = element_blank(),
  panel.border = element_rect(color='black', fill=NA, size=1),
  plot.title=element_blank()
)

pdf(paste0(fig_dir, 'ODC_module_preservation_summary_Morabito.pdf'), width=3, height=3)
p
dev.off()

```


Correlation of Seurat module score / UCell score and MEs:

1. Subset each cell type
2. Project modules (no harmony)
3. Compute Seurat score
4. Save the subset object

```{r eval=FALSE}

cur_group <- 'ASC'

groups <- c('ASC', 'EX', 'INH', 'MG', 'ODC', 'OPC')
for(cur_group in groups){

  seurat_subset <- subset(seurat_obj, cell_type == cur_group)

  seurat_subset <- ProjectModules(
    seurat_obj = seurat_subset,
    seurat_ref = seurat_obj,
    wgcna_name = cur_group,
    wgcna_name_proj=paste0(cur_group, '_subset'),
    assay="RNA"
  )
  seurat_subset <- ModuleExprScore(seurat_subset, n_genes = Inf)

  scores <- GetModuleScores(seurat_subset)
  MEs <- GetMEs(seurat_subset)
  modules <- GetModules(seurat_subset)
  mods <- levels(modules$module); mods <- mods[mods != 'grey']

  # make a scatter plot of MEs vs Scores
  scatter_list <- list()
  for(cur_mod in mods){
    print(cur_mod)
    cur_color <- subset(modules, module == cur_mod) %>% .$color %>% unique()
    plot_df <- data.frame(
      ME = MEs[,cur_mod],
      score = scores[,cur_mod],
      module = cur_mod,
      color = cur_color
    )

    p <- ggplot(plot_df, aes(x=ME, y=score)) +
    geom_hline(yintercept=0, linetype='dashed', color='grey') +
    geom_vline(xintercept=0, linetype='dashed', color='grey') +
    ggrastr::rasterise(
      geom_point(color=cur_color),
      dpi=400,
      scale=0.5
    ) +
    geom_smooth(method='lm', color='black', fill='lightgrey') +
    stat_cor(method='pearson') +
    ggtitle(cur_mod) +
    theme(
      plot.title = element_text(hjust=0.5, face='bold'),
      panel.border = element_rect(color='black', fill=NA, size=1),
      axis.line.y = element_blank(),
      axis.line.x = element_blank(),
      axis.title.y = element_blank(),
      axis.title.x = element_blank(),
      plot.margin = margin(0,0,0,0)
    ) +
    NoLegend()

    scatter_list[[cur_mod]] <- p

  }

  pdf(paste0(fig_dir, 'ME_score_correlation_', cur_group,'.pdf'), width=12, height=12)
  print(wrap_plots(scatter_list, ncol=5, nrow=5))
  dev.off()
}


```


Differential module eigengene testing to compare cell types

```{r eval=FALSE}

DMEs <- do.call(rbind, lapply(groups, function(cur_group){
  dmes <- FindAllDMEs(
    seurat_obj,
    group.by = 'cell_type',
    wgcna_name=cur_group
  )
  dmes$group
}))

write.csv(DMEs, row.names=FALSE, quote=FALSE, file=paste0(data_dir, 'brain_DMEs.csv'))

```
