

Using the Zhou et al hdWGCNA analysis from Figure 1 of the paper, we need to
evaluate the performance of the method as suggested by the peer reviewers.

```{r eval=FALSE}


library(Seurat)
library(tidyverse)
library(cowplot)
library(patchwork)
library(WGCNA)

library(Matrix)
library(viridis)
library(harmony)
library(RColorBrewer)
library(ggpubr)
library(tictoc)
library(RColorBrewer)
library(Hmisc)
library(corrplot)
library(enrichR)
library(GeneOverlap)
library(grid)
library(gridExtra)
library(igraph)
library(ggrepel)
library(hdWGCNA)
library(magrittr)
enableWGCNAThreads(nThreads = 8)
theme_set(theme_cowplot())
set.seed(12345)


# spatial plotting functions
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

#devtools::install_github('smorabit/hdWGCNA', ref='dev')
#library(hdWGCNA)

setwd('/dfs7/swaruplab/smorabit/analysis/scWGCNA/')

# output directories
data_dir <- "data/"
fig_dir <- 'figures/'

# re-load scWGCNA dataset:
load(file='data/Zhou_color_scheme.rda')


# reload process ed with all cell types
# seurat_obj <-  readRDS(file=paste0(data_dir, 'Zhou_scWGCNA_all_celltypes.rds'))


load('/dfs7/swaruplab/smorabit/analysis/AD_NucSeq_2019/batch_correction/liger/update/celltype-analysis/data/color_scheme.rda')
cp <- unlist(color_scheme_snRNA_celltype)
cp[names(cp) == 'EX'] <- 'turquoise'


setwd('/dfs7/swaruplab/smorabit/analysis/scWGCNA/evaluate')

seurat_obj <- readRDS(file='/dfs7/swaruplab/smorabit/analysis/scWGCNA/brain_iterative/data/Zhou_scWGCNA_all_celltypes.rds')


```

Run EGAD tutorial

https://github.com/sarbal/EGAD

```{r eval=FALSE}

# Load EGAD and the data files
library(EGAD)
data(biogrid)
data(GO.human)

# Or you can load EGADlite here too (https://github.com/sarbal/EGADLite):
# load("EGADlite.RData")
# download the data folder into your directory and run
# load("data/biogrid.RData")
# load("data/GO.human.RData")

# Make your gene list and the network
genelist <- make_genelist(biogrid)
gene_network <- make_gene_network(biogrid,genelist)

# Store your annotation matrix
goterms <- unique(GO.human[,3])
annotations <- make_annotations(GO.human[,c(2,3)],genelist,goterms)

# Run GBA
GO_groups_voted <- run_GBA(gene_network, annotations)

# neighbor voting AUROCs
auc_GO_nv = GO_groups_voted[[1]][,1]
median(auc_GO_nv)

# node degree AUCs
auc_GO_nd = GO_groups_voted[[1]][,3]
median(auc_GO_nd)

length(goterms)
length(genelist)

dim(annotations)

# all the genes and only some of the GO terms
dim(GO_groups_voted[[2]])

df <- data.frame(
  nv_auc = auc_GO_nv,
  nd_auc = auc_GO_nd
)



p1 <- ggplot(df, aes(y = nv_auc, x=1)) +
  geom_violin() +
  geom_boxplot(notch=TRUE, width=0.25, fill='white', outlier.size=0) +
  ylim(c(0,1)) +
  xlab('') + ylab('AUC') +
  ggtitle('Neighbor voting') +
  theme(
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    panel.border =element_rect(color='black', fill=NULL, size=1),
    plot.title = element_text(vjust=-0.5, hjust=0.5, face='plain'),
    panel.grid.major.y = element_line(color='lightgrey', size=0.5)
  )


pdf(paste0(fig_dir, 'test_dens_example.pdf'), width=2, height=3)
p1
dev.off()


```

Try to run EGAD on our data

Need to loop over each cell type

```{r eval=FALSE}

library(EGAD)

groups <- c('ASC', 'EX', 'INH', 'MG', 'OPC', 'ODC')

ensembl <- read.delim('/dfs7/swaruplab/smorabit/resources/ensembl_GO.txt', sep='\t', header=1)

# load the genes file
gene_table <- read.table('/dfs7/swaruplab/shared_lab/cross-disorder/count_matrices/AD_Zhou_2020/C1/counts_unfiltered/genes.tsv')
gene_table$ID <- do.call(rbind, strsplit(gene_table$V1, '[.]'))[,1]


plot_df <- data.frame()
auc_df <- data.frame()
for(cur_group in groups){

  print(cur_group)

  # get network info for this group
  TOM <- GetTOM(seurat_obj, cur_group)
  modules <- GetModules(seurat_obj, cur_group)

  # get a list of unique genes
  genelist <- unique(modules$gene_name)
  genelist <- genelist[(genelist %in% gene_table$V2)]

  # subset TOM to just these genes
  TOM <- TOM[genelist,genelist]

  # get IDs for these genes
  ix <- match(genelist, gene_table$V2)
  gene_ids <- gene_table$ID[ix]

  # rename TOM genes to gene ids
  colnames(TOM) <- gene_ids
  rownames(TOM) <- gene_ids

  # set up annotations
  goterms <- unique(ensembl$GO.term.accession)
  go_table <- ensembl[,c(1,4)] %>% subset(GO.term.accession != '')
  annotations <- make_annotations(go_table, gene_ids, goterms)

  print(dim(annotations))
  length(gene_ids)
  length(goterms)

  # Run GBA
  GO_groups_voted <- run_GBA(TOM, annotations, min=10, max=Inf)

  # neighbor voting AUROCs
  auc_GO_nv = GO_groups_voted[[1]][,1]

  dim(GO_groups_voted[[2]])

  # node degree AUCs
  auc_GO_nd = GO_groups_voted[[1]][,3]

  # multifunctionality
  multifunc_assessment <- calculate_multifunc(annotations)
  auc_mf <- auc_multifunc(annotations, multifunc_assessment[,4])
  df <- data.frame(
    auc = auc_mf,
    annotation = colnames(annotations)
  )
  df$celltype <- cur_group
  auc_df <- rbind(auc_df, df)

  # neighbor voting df
  df <- data.frame(
    nv_auc = auc_GO_nv,
    nd_auc = auc_GO_nd
  )
  df$celltype <- cur_group
  plot_df <- rbind(plot_df, df)

}



auc_df %<>% subset(!is.nan(auc))

t1 <- subset(auc_df, celltype == 'EX')
t2 <- subset(auc_df, celltype == 'INH')

all.equal(t1$annotation, t2$annotation)
all.equal(t1$auc, t2$auc)

plot_df %>% group_by(celltype) %>% summarise(Mean = median(nv_auc))

subset(plot_df, celltype == 'ASC') %>% .$nv_auc %>% quantile

# annotations <- c(rep(0,10))
# annotations[c(1,3,5)] <- 1
# optimallist <- 10:1
# aurocs_mf <- auc_multifunc(annotations, optimallist)



p1 <- ggplot(plot_df, aes(y = nv_auc, x=reorder(celltype, nv_auc, median), fill=celltype)) +
  geom_violin() +
  geom_boxplot(notch=TRUE, width=0.25, fill='white', outlier.size=0) +
  ylim(c(0,1)) +
  xlab('') + ylab('AUC') +
  scale_fill_manual(values=cp) +
  RotatedAxis() +
  NoLegend() +
  ggtitle('Neighbor voting') +
  theme(
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    panel.border =element_rect(color='black', fill=NULL, size=1),
    plot.title = element_text(vjust=-0.5, hjust=0.5, face='plain'),
    panel.grid.major.y = element_line(color='lightgrey', size=0.5)
  )

auc_df$celltype <- factor(
  as.character(auc_df$celltype),
  levels = c('ODC', 'EX', 'OPC', 'ASC', 'MG', 'INH')
)

p2 <- ggplot(auc_df, aes(y = auc, x=celltype, fill=celltype)) +
  geom_violin() +
  geom_boxplot(notch=TRUE, width=0.25, fill='white', outlier.size=0) +
  ylim(c(0,1)) +
  xlab('') + ylab('AUC') +
  scale_fill_manual(values=cp) +
  RotatedAxis() +
  NoLegend() +
  ggtitle('Multifunctionality') +
  theme(
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    panel.border = element_rect(color='black', fill=NULL, size=1),
    plot.title = element_text(vjust=-0.5, hjust=0.5, face='plain'),
    panel.grid.major.y = element_line(color='lightgrey', size=0.5)
  )



p3 <- ggplot(plot_df, aes(y = nd_auc, x=reorder(celltype, nd_auc, mean), fill=celltype)) +
  geom_violin() +
  geom_boxplot(notch=TRUE, width=0.25, fill='white', outlier.size=0) +
  ylim(c(0,1)) +
  xlab('') + ylab('AUC') +
  scale_fill_manual(values=cp) +
  RotatedAxis() +
  NoLegend() +
  ggtitle('Node degree') +
  theme(
    axis.line.y = element_blank(),
    axis.line.x = element_blank(),
    panel.border =element_rect(color='black', fill=NULL, size=1),
    plot.title = element_text(vjust=-0.5, hjust=0.5, face='plain'),
    panel.grid.major.y = element_line(color='lightgrey', size=0.5)
  )



# p2 <- ggplot(df, aes(y = nd_auc, x=1)) +
#   geom_violin() +
#   ylim(c(0,1)) +
#   xlab('')



pdf(paste0(fig_dir, 'test_dens2.pdf'), width=3, height=3)
p1
dev.off()

pdf(paste0(fig_dir, 'test_multifunc2.pdf'), width=3, height=3)
p2
dev.off()

pdf(paste0(fig_dir, 'test_degree2.pdf'), width=3, height=3)
p3
dev.off()


```

Testing code for tracking runtime and memory of different functions

1. MetacellsByGroups
2. MetaspotsByGroups
3. ConstructNetwork
4. ModuleEigengenes
5. ModuleConnectivity

```{r eval=FALSE}


seurat_asd <- readRDS(file='/dfs7/swaruplab/smorabit/analysis/scWGCNA/ASD_case_study/data/Velmeshev_2019_hdWGCNA.rds')
table(seurat_asd$cell_type)

seurat_obj <- seurat_asd %>% subset(cell_type %in% c('INH', 'EX'))
seurat_obj$cell_group <- 'Neuron'

library(tictoc)

mem_start <- gc()

sample_sizes <- c(1000, 5000, 10000, 15000, 20000, 25000, 30000, 35000, 40000, 45000, 50000)

runtime_df <- data.frame()
for(n_sample in sample_sizes){

  ################################################################################
  # Setup the data
  ################################################################################
  print(n_sample)
  # downsample to 1,000 cells
  #n_sample <- 1000
  #seurat_sample <- subset(seurat_obj, cell_type == 'ODC')

  cells_sample <- sample(colnames(seurat_obj), n_sample)
  seurat_sample <- seurat_obj[,cells_sample]
  print(dim(seurat_sample))

  seurat_sample <- SetupForWGCNA(
    seurat_sample,
    gene_select = "fraction",
    fraction = 0.05,
    wgcna_name = 'test'
  )

  ################################################################################
  # MetaceLlsByGroups
  ################################################################################
  print('MetacellsByGroups')

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- MetacellsByGroups(
    seurat_sample,
    group.by = c("cell_group", "Sample"),
    k = 10,
    max_shared = 5,
    reduction = 'harmony',
    ident.group = 'cell_group',
    min_cells = 20
  )
  mcell_time <- toc()
  mcell_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt


  ################################################################################
  # SetDatExpr & TestSoftPowers (not timed)
  ################################################################################
  print('SetDatExpr')

  seurat_sample <- NormalizeMetacells(seurat_sample)

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- SetDatExpr(
    seurat_sample,
    group.by='cell_group',
    group_name = unique(seurat_sample$cell_group),
    use_metacells=TRUE,
    slot = 'data',
    assay = 'RNA'
  )
  sd_time <- toc()
  sd_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt

  seurat_sample <- TestSoftPowers(seurat_sample, setDatExpr = FALSE)


  ################################################################################
  # ConstructNetwork
  ################################################################################
  print('ConstructNetwork')

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- ConstructNetwork(
    seurat_sample,
    setDatExpr=FALSE,
    tom_name = paste0('test', n_sample),
    overwrite_tom=TRUE
  )
  cn_time <- toc()
  cn_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt



  ################################################################################
  # ModuleEigengenes with harmony
  ################################################################################
  print('ModuleEigengenes + harmony')

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- ModuleEigengenes(
    seurat_sample,
    group.by.vars="Sample"
  )
  meh_time <- toc()
  meh_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt

  ################################################################################
  # ModuleEigengenes
  ################################################################################
  print('ModuleEigengenes')

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- ModuleEigengenes(seurat_sample)
  me_time <- toc()
  me_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt

  ################################################################################
  # ModuleConnectivity
  ################################################################################
  print('ModuleConnectivity')

  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_sample <- ModuleConnectivity(
    seurat_sample,
    group.by='cell_group',
    group_name = as.character(unique(seurat_sample$cell_group))
  )
  mc_time <- toc()
  mc_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt



  ################################################################################
  # Consolidate runtime information
  ################################################################################

  modules <- GetModules(seurat_sample)
  n_mods <- length(unique(modules$module))

  run_df <- data.frame(
    'func' = c('MetacellsByGroups', 'SetDatExpr', 'ConstructNetwork', 'ModuleEigengenes', 'ModuleEigengenesHarmony', 'ModuleConnectivity'),
    'runtime' = c(
      as.numeric(mcell_time$toc - mcell_time$tic),
      as.numeric(sd_time$toc - sd_time$tic),
      as.numeric(cn_time$toc - cn_time$tic),
      as.numeric(me_time$toc - me_time$tic),
      as.numeric(meh_time$toc - meh_time$tic),
      as.numeric(mc_time$toc - mc_time$tic)
    ),
    'memory' = c(
      mcell_mem, sd_mem, cn_mem, me_mem, meh_mem, mc_mem
    )
  )
  # run_df$memstr <- R.utils::hsize(run_df$memory)
  run_df$sample <- n_sample

  runtime_df <- rbind(runtime_df, run_df)

  rm(seurat_sample)
  gc()

}

write.csv(runtime_df, file='data/runtime_summary2.csv')

################################################################################
# Plot runtime results
################################################################################


# memory size of object
# test <- GetMEs(seurat_sample)
# object.size(GetMEs(seurat_sample)) / 1024 / 1024

# set factor levels for function
runtime_df$func <- ifelse(
  as.character(runtime_df$func) == 'ModuleEigengenesHarmony',
  'ME + Harmony',
  as.character(runtime_df$func)
)

# compute totals
total_df <- runtime_df %>% group_by(sample) %>% summarise(runtime = sum(runtime), memory = sum(memory))
total_df$memstr <- R.utils::hsize(total_df$memory)
total_df$func <- 'Total'
total_df <- total_df[,colnames(runtime_df)]
runtime_df <- rbind(runtime_df, total_df)

runtime_df$func <- factor(
  as.character(runtime_df$func),
  levels = c('MetacellsByGroups', 'SetDatExpr', 'ConstructNetwork', 'ModuleEigengenes', 'ME + Harmony', 'ModuleConnectivity', 'Total'),
)

runtime_df <- subset(runtime_df, func != 'Total') %>% mutate(func = droplevels(func))


p <- ggplot(runtime_df, aes(x=sample, y=memory/1024)) +
  geom_point(color = 'darkgoldenrod3') +
  geom_line(color = 'darkgoldenrod3') +
  RotatedAxis() +
  ylab('Memory upper bound (Gb)') +
  xlab(expression(N[cells])) +
  scale_x_continuous(label=comma) +
  theme(
    panel.grid.major =element_line(color='lightgrey', size=0.5),
    panel.border = element_rect(size=1, fill=NA, color='black'),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    strip.text.x = element_text(size=9)
  )

pdf(paste0(fig_dir, 'test_memory3.pdf'), width=5, height=3)
p + facet_wrap(~func, ncol=3)
# p + facet_wrap(~func, ncol=4, scales='free_y')
dev.off()



p <- ggplot(runtime_df, aes(x=sample, y=runtime / 60)) +
  geom_point(color='darkorchid3') +
  geom_line(color='darkorchid3') +
  RotatedAxis() +
  ylab('Runtime (minutes)') +
  xlab(expression(N[cells])) +
  scale_x_continuous(label=comma) +
  theme(
    panel.grid.major =element_line(color='lightgrey', size=0.5),
    panel.border = element_rect(size=1, fill=NA, color='black'),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    strip.text.x = element_text(size=9)
  )

pdf(paste0(fig_dir, 'test_runtime3.pdf'), width=5, height=3)
# p + facet_wrap(~func, ncol=4, scales='free_y')
p + facet_wrap(~func, ncol=3)
dev.off()


```

Additional tests:

1. Test the runtime of MetacellsByGroups changing the number of target metacells
2. Runtime of ModuleConnectivity with downsampling
   1. RRHO or spearman correlation of the resulting kMEs
3. Runtime of ConstructNetwork with different numbers of genes as input

```{r eval=FALSE}

################################################################################
# Test Metacells
################################################################################

target_metacells <- c(50, 100, 200, 500, 1000)

mcell_run_df <- data.frame()

for(cur_target in target_metacells){
  print(cur_target)
  tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
  tic()
  seurat_obj <- MetacellsByGroups(
    seurat_obj,
    group.by = c("cell_type", "Sample"),
    k = 25,
    max_shared = 12,
    reduction = 'harmony',
    ident.group = 'cell_type',
    min_cells = 50,
    target_metacells = cur_target
  )
  mcell_time <- toc()
  mcell_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt


  cur_run_df <- data.frame(
    'runtime' = as.numeric(mcell_time$toc - mcell_time$tic),
    'memory' = mcell_mem,
    'target' = cur_target,
    'n_metacells' = ncol(GetMetacellObject(seurat_obj))
  )

  mcell_run_df <- rbind(mcell_run_df, cur_run_df)

  seurat_obj <- SetMetacellObject(seurat_obj, NULL)
  gc()

}




write.csv(mcell_run_df, file='data/mcell_runtime_summary.csv')

################################################################################
# Plot runtime results
################################################################################

p <- ggplot(mcell_run_df, aes(x=target, y=memory/1024)) +
  geom_point() +
  geom_line() +
  RotatedAxis() +
  ylab('Memory upper bound (Gb)') +
  xlab(expression(N[target])) +
  scale_x_continuous(label=comma) +
  theme(
    panel.grid.major =element_line(color='lightgrey', size=0.5),
    panel.border = element_rect(size=1, fill=NA, color='black'),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    strip.text.x = element_text(size=9)
  )

pdf(paste0(fig_dir, 'test_memory_mcell.pdf'), width=2.5, height=3)
p
# p + facet_wrap(~func, ncol=4, scales='free_y')
dev.off()



p <- ggplot(mcell_run_df, aes(x=target, y=runtime / 60)) +
  geom_point() +
  geom_line() +
  RotatedAxis() +
  ylab('Runtime (minutes)') +
  xlab(expression(N[target])) +
  scale_x_continuous(label=comma) +
  theme(
    panel.grid.major =element_line(color='lightgrey', size=0.5),
    panel.border = element_rect(size=1, fill=NA, color='black'),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    strip.text.x = element_text(size=9),
    axis.title.y = element_blank()
  )

pdf(paste0(fig_dir, 'test_runtime_mcell.pdf'), width=2.5, height=2)
# p + facet_wrap(~func, ncol=4, scales='free_y')
p
dev.off()


```

Test re-writing module Connectivity using fastCor

```{r eval=FALSE}

library(HiClimR)
library(qlcMatrix)
library(tictoc)

# get modules and MEs
wgcna_name <- 'INH'
group.by <- 'cell_type'; group_name <- 'INH'

modules <- GetModules(seurat_obj, wgcna_name)
MEs <- GetMEs(seurat_obj, harmonized=TRUE, wgcna_name)
cells.use <- seurat_obj@meta.data %>% subset(get(group.by) %in% group_name) %>% rownames
genes_use <- GetWGCNAGenes(seurat_obj, wgcna_name)


# get expression data:
exp_mat <- GetAssayData(
  seurat_obj,
  assay='RNA',
  slot='data'
)[genes_use,cells.use]
datExpr <- t(as.matrix(exp_mat))

#  transpose
exp_mat <- t(exp_mat)

MEs <- MEs[cells.use,]



# correlate

tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
tic()
test <- qlcMatrix::corSparse(
  X = exp_mat,
  Y = as.matrix(MEs)
)
rownames(test) <- genes_use
colnames(test) <- colnames(MEs)
colnames(test) <- paste0("kME_", colnames(test))

cor_time <- toc()
cor_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt
cor_time$toc - cor_time$tic
cor_mem



# correlate w/ wgcna function

tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
tic()

kMEs <- WGCNA::signedKME(
  datExpr,
  MEs,
  outputColumnName = "kME",
  corFnc = 'bicor',
  corOptions = "use='p'"
)
colnames(kMEs) <- colnames(MEs)
colnames(kMEs) <- paste0("kME_", colnames(kMEs))

kme_time <- toc()
kme_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt
kme_time$toc - kme_time$tic
kme_mem



tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
tic()

seurat_obj <- ModuleConnectivity(
  seurat_obj,
  group.by = 'cell_type',
  group_name = 'INH',
  wgcna_name = wgcna_name
)

mc_time <- toc()
mc_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt
mc_time$toc - mc_time$tic
mc_mem



tt <- sum(.Internal(gc(FALSE, TRUE, TRUE))[13:14])
tic()

seurat_obj <- ModuleConnectivity(
  seurat_obj,
  group.by = 'cell_type',
  group_name = 'INH',
  sparse = FALSE,
  wgcna_name = wgcna_name
)

mco_time <- toc()
mco_mem <- sum(.Internal(gc(FALSE, FALSE, TRUE))[13:14]) - tt
mco_time$toc - mco_time$tic
mco_mem




# correlation of the kME vs sparseCor:
df <- reshape2::melt(as.matrix(kMEs))
df2 <- reshape2::melt(test)
all.equal(df$Var1, df2$Var2)

plot_df <- data.frame(
  kME = df$value,
  sparseCor = df2$value,
  gene = df$Var1,
  module = do.call(rbind, strsplit(as.character(df$Var2), '_'))[,2]
)
plot_df$module <- factor(
  as.character(plot_df$module),
    levels=levels(modules$module)
)

mod_colors <- modules %>% select(c(module, color)) %>% distinct() %>% arrange(module)
mod_cp <- mod_colors$color; names(mod_cp) <- mod_colors$module

p <- plot_df %>% ggplot(aes(x = kME, y = sparseCor, color=module)) +
  ggrastr::rasterise(geom_point(), dpi=300) +
  geom_smooth(method = 'lm', color='grey5') +
  ggpubr::stat_cor(label.y.npc="top", label.x.npc = "left", method = "pearson", size=2.5, color='black') +
  scale_color_manual(values=mod_cp) +
  scale_x_continuous(limits=c(-1,1)) +
  scale_y_continuous(limits=c(-1,1)) +
  theme(
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    panel.border = element_rect(size=1, fill=NA, color='black'),
    panel.grid.major = element_line(size=0.5, color='lightgrey')
  ) + NoLegend()

pdf(paste0(fig_dir, 'test_kme_sparsecor.pdf'), width=10, height=8)
p + facet_wrap(~module, ncol=5, scales='free')
dev.off()





# plot kMEs
p <- PlotKMEs(seurat_obj, ncol=5, wgcna_name = wgcna_name)

pdf(paste0(fig_dir, 'test_kmes.pdf'), width=10, height=8)
p
dev.off()


################################################################################
# Reassign modules kMEs with negative values
################################################################################

ReassignModules <- function(
  seurat_obj,
  harmonized=TRUE,
  features=NULL,
  new_modules=NULL,
  ignore=FALSE,
  wgcna_name=NULL
){

  if(is.null(wgcna_name)){wgcna_name <- seurat_obj@misc$active_wgcna}

  # get modules and MEs from seurat object
  modules <- GetModules(seurat_obj, wgcna_name)
  MEs <- GetMEs(seurat_obj, harmonized, wgcna_name)
  mod_colors <- dplyr::select(modules, c(module, color)) %>% distinct()
  mod_cp <- mod_colors$color; names(mod_cp) <- as.character(mod_colors$module)
  mods <- levels(modules$module); mods <- mods[mods != 'grey']
  genes_use <- GetWGCNAGenes(seurat_obj, wgcna_name)


  if(!is.null(features) & !is.null(new_modules)){

    ##############################
    # Manual reassignment
    ##############################

    # check validity of input features
    if(!all(features %in% genes_use)){
      stop('Some features are not found in GetWGCNAGenes(seurat_obj).')
    }

    # check validity of modules
    if(!all(new_modules %in% levels(modules$module))){
      stop('Some module names in new_modules are invalid. Features must be reassigned to existing modules found in GetModules(seurat_obj)')
    }

    # get original modules
    orig_mods <- subset(modules, gene_name %in% features) %>% .$module %>% as.character()
    if(!all(orig_mods == 'grey') & !ignore){
      stop('Attempting to reassign non-grey genes to new modules. Proceed with caution. If you wish to reassign these genes, re-run this function and set ignore=TRUE')
    }

    # set up table for new module assignments
    reassign_df <- data.frame(
      gene_name = as.character(features),
      module = as.character(new_modules)
    )
    reassign_df$module <- factor(as.character(reassign_df$module), levels = levels(modules$module))

    # new colors:
    reassign_df$color <- as.character(mod_cp[as.character(reassign_df$module)])

    # reassign modules and colors
    modules[reassign_df$gene_name,'module'] <- reassign_df$module
    modules[reassign_df$gene_name,'color'] <- reassign_df$color

  } else{


    ##############################
    # reassignment by kME
    ##############################

    # get genes with negative kME in their assigned module:
    neg_df <- do.call(rbind, lapply(mods, function(cur_mod){
      cur <- subset(modules, module == cur_mod)
      cur <- cur[,c('gene_name', 'module', paste0('kME_', cur_mod))]
      names(cur)[3] <- 'kME'
      cur %>% subset(kME < 0)
    }))
    if(nrow(neg_df) == 0){
      stop('No genes to reassign, all kMEs of assigned modules are greater than 0.')
    }
    rownames(neg_df) <- 1:nrow(neg_df)

    print(dim(neg_df))

    # get just the kME values from the modules table
    kMEs <- modules[,4:ncol(modules)]
    kMEs <- kMEs[,colnames(kMEs) != "kME_grey"]

    # for each gene with negative kME values in the assigned module,
    # identify the module that had the highest kME
    reassigned <- sapply(neg_df$gene_name, function(cur_gene){
      cur_kMEs <- kMEs[cur_gene,]
      max_kME <- max(cur_kMEs)
      if(max_kME < 0){
        return('kME_grey')
      }
      colnames(kMEs)[which(cur_kMEs == max_kME)]
    })

    # add the reassigned modules to the modules table
    reassigned <- do.call(rbind, strsplit(reassigned, 'kME_'))[,2]
    reassigned <- factor(as.character(reassigned), levels=levels(modules$module))

    # new colors:
    reassigned_color <- as.character(mod_cp[as.character(reassigned)])

    # reassign modules and colors
    modules[neg_df$gene_name,'module'] <- reassigned
    modules[neg_df$gene_name,'module'] <- reassigned_colors

  }

  # set the modules table
  seurat_obj <- SetModules(seurat_obj, modules, wgcna_name)
  seurat_obj
}

old_modules <- GetModules(seurat_obj, wgcna_name)

# re-assign based on kME
seurat_obj <- ReassignModules(
  seurat_obj, wgcna_name=wgcna_name
)

# re-assign manually
seurat_obj <- ReassignModules(
  seurat_obj, wgcna_name=wgcna_name,
   features=c('SLC17A7', 'GFAP'), new_modules = c('grey', 'INH-M7'),
   ignore=FALSE
)

################################################################################
# Plot Reassigned genes with ggsankey
################################################################################

library(ggsankey)

mod_colors <- modules %>% select(c(module, color)) %>% distinct() %>% arrange(module)
mod_cp <- mod_colors$color; names(mod_cp) <- mod_colors$module


df <- mtcars %>%
  make_long(cyl, vs, am, gear, carb)

df <- neg_df %>%
  make_long(module, reassigned)

p <- ggplot(df, aes(x = x,
               next_x = next_x,
               node = node,
               next_node = next_node,
               fill = factor(node),
               color = factor(node)
               label = node)) +
  geom_sankey(flow.alpha=0.75) +
  geom_sankey_label(size = 3.5, color = 1, fill = "white") +
  theme_sankey(base_size = 16) +
  scale_fill_manual(values=mod_cp) +
  NoLegend() +
  xlab('')

pdf(paste0(fig_dir, 'test_kmes_reassigned.pdf'), width=6, height=8)
p
dev.off()


```
