
```{r eval=FALSE}

library(Seurat)
library(Matrix)
library(tidyverse)
library(cowplot)
library(ggpubr)
library(harmony)
library(patchwork)
library(WGCNA)
library(viridis)
library(RColorBrewer)
library(ggpubr)
library(tictoc)
library(Hmisc)
library(igraph)
library(ggrepel)
library(hdWGCNA)
enableWGCNAThreads(nThreads = 8)
set.seed(12345)
theme_set(theme_cowplot())

# load custom plotting script
source("/pub/smorabit/hdWGCNA/bin/spatial_functions.R")
source('/dfs7/swaruplab/smorabit/analysis/scWGCNA/bin/spatial_functions.R')

scripts <- dir("/dfs7/swaruplab/smorabit/analysis/galilean/bin/")
for(script in scripts){
  source(paste0("/dfs7/swaruplab/smorabit/analysis/galilean/bin/", script))
}

# output directories
fig_dir <- 'figures/'
data_dir <- 'data/'


```

Set up ScISOrSeq dataset

```{r eval=FALSE}

# load ScISOrSeq  dataset:
iso_X <- Seurat::Read10X("/pub/smorabit/hdWGCNA/isoseq/data/dataSharing_countsPerCell/isoformByCellMatrix/")
iso_table <- read.delim("/pub/smorabit/hdWGCNA/isoseq/data/dataSharing_countsPerCell/IsoID_TranscriptID_TSS_Intron_PolyA", sep='\t', header=FALSE)
ct_table <- read.delim("/pub/smorabit/hdWGCNA/isoseq/data/dataSharing_countsPerCell/Barcode_Celltype_Assignments", sep='\t', header=FALSE)

ct_table <- dplyr::rename(ct_table, c(barcode=V1, group=V2))
ct_table$celltype <- do.call(rbind, strsplit(ct_table$group, '_'))[,2]
rownames(ct_table) <- ct_table$barcode

# load illumina data:
rep1_X <- Seurat::Read10X("/pub/smorabit/hdWGCNA/isoseq/data/illumina/rep1/")
rep2_X <- Seurat::Read10X("/pub/smorabit/hdWGCNA/isoseq/data/illumina/rep2/")

colnames(rep1_X) <- do.call(rbind, strsplit(colnames(rep1_X), '-'))[,1]
colnames(rep2_X) <- do.call(rbind, strsplit(colnames(rep2_X), '-'))[,1]

s1 <- CreateSeuratObject(rep1_X); s1$barcode <- colnames(s1); s1$rep <- "rep1"
s2 <- CreateSeuratObject(rep2_X); s2$barcode <- colnames(s2); s2$rep <- "rep2"

seurat_obj <- merge(s1, s2)

# get rid of duplicated barcodes:
single_barcodes <- table(seurat_obj$barcode) == 1
single_barcodes <- single_barcodes[seurat_obj$barcode]
seurat_obj <- seurat_obj[,single_barcodes]

seurat_obj[["percent.mt"]] <- PercentageFeatureSet(seurat_obj, pattern = "^mt-")

# which cells do we have long read data for?
seurat_obj$LR <- seurat_obj$barcode %in% colnames(iso_X)
seurat_obj <- subset(seurat_obj, LR)

# subset to have the same cells as the seurat obj:
iso_X <- iso_X[,seurat_obj$barcode]
all.equal(colnames(iso_X), as.character(seurat_obj$barcode))

# make column names the same as the seurat obj
colnames(iso_X) <- colnames(seurat_obj)

# add the iso assay to the seurat obj
iso_assay <- Seurat::CreateAssayObject(counts = iso_X)
seurat_obj[["iso"]] <- iso_assay

# add celltype information:
head(ct_table)
ct_table <- ct_table[seurat_obj$barcode,]
seurat_obj$celltype <- ct_table$celltype

saveRDS(seurat_obj, paste0(data_dir, 'sniso_seurat_unprocessed.rds'))

```

Dimensionality reduction on RNA assay

```{r eval=FALSE}

# re-load unprocessed dataset:
seurat_obj <- readRDS(paste0(data_dir, 'sniso_seurat_unprocessed.rds'))
DefaultAssay(seurat_obj) <- 'RNA'

# standard Seurat processing:
seurat_obj <- seurat_obj %>%
  NormalizeData() %>%
  FindVariableFeatures() %>%
  ScaleData() %>%
  RunPCA()

# seurat_obj <- RunHarmony(seurat_obj, 'rep')
seurat_obj <- RunUMAP(seurat_obj, dims=1:30, min.dist=0.3)
saveRDS(seurat_obj, paste0(data_dir, 'sniso_seurat_processed.rds'))


# plot the UMAP
p <- PlotEmbedding(
  seurat_obj,
  group.by = 'annotation',
  label=TRUE,
  raster=TRUE,
  raster_dpi = 200,
  point_size=0.25,
  raster_scale=0.5,
  plot_under=TRUE,
  plot_theme = umap_theme()
)

pdf(paste0(fig_dir, 'umap_anno.pdf'), width=7, height=7)
p
dev.off()


```

Run isoform co-expression network analysis in the radial glia lineage with hdWGCNA

```{r eval=FALSE}

# re-load unprocessed dataset:
seurat_obj <- readRDS(paste0(data_dir, 'sniso_seurat_processed.rds'))

# add broad annotation
ct_df <- read.csv('~/pub/hdWGCNA/isoseq/data/sniso_annotation.csv')
ix <- match(seurat_obj$celltype, ct_df$celltype)
seurat_obj$annotation <- ct_df$annotation[ix]

# switch to isoform assay
DefaultAssay(seurat_obj) <- 'iso'

keep <- c('NIPCs', 'ASC', 'RGL', 'Ependymal')
seurat_obj <- seurat_obj[,seurat_obj$annotation %in%keep]


seurat_obj <- SetupForWGCNA(
  seurat_obj,
  gene_select = "fraction",
  fraction = 0.01,
  group.by = 'annotation',
  wgcna_name = "test"
)
length(GetWGCNAGenes(seurat_obj))

# construct metacells:
seurat_obj <- MetacellsByGroups(
  seurat_obj = seurat_obj,
  group.by = c("annotation"),
  k = 30,
  max_shared=20,
  min_cells=50,
  reduction = 'pca',
  ident.group = 'annotation',
  assay = 'iso',
  mode = 'sum'
)

seurat_obj <- NormalizeMetacells(seurat_obj)

# setup expression matrix
seurat_obj <- SetDatExpr(
  seurat_obj,
  group.by='annotation',
  group_name = unique(seurat_obj$annotation),
  use_metacells=TRUE
)

seurat_obj <- TestSoftPowers(seurat_obj)

# plot the results:
plot_list <- PlotSoftPowers(seurat_obj)

# assemble with patchwork
pdf(paste0(fig_dir, 'RGL_softpower.pdf'), width=12, height=9)
wrap_plots(plot_list, ncol=2)
dev.off()

# construct wgcna network:
seurat_obj <- ConstructNetwork(
  seurat_obj,
  soft_power=5,
  setDatExpr=FALSE,
  detectCutHeight=0.995,
  mergeCutHeight=0.5
)

# plot the dendrogram
pdf(paste0(fig_dir, "RGL_dendro.pdf"),height=3, width=6)
PlotDendrogram(seurat_obj, main='RGL hdWGCNA Dendrogram')
dev.off()

# compute MEs and module connectivity:
seurat_obj <- ModuleEigengenes(seurat_obj,)
seurat_obj <- ModuleConnectivity(seurat_obj)


################################################################################
# reset module names and colors
################################################################################

# run RenameModules
seurat_obj <- ResetModuleNames(
  seurat_obj,
  new_name = "RGL-M"
)
print(names(GetModules(seurat_obj)))


library(MetBrewer)
new_colors <- paste0(met.brewer("Renoir", n=11, type='discrete'))
new_colors <- sample(new_colors, 11 )
seurat_obj <- ResetModuleColors(seurat_obj, new_colors)



################################################################################
# ME feature plots
################################################################################

plot_list <- ModuleFeaturePlot(
  seurat_obj,
  order=TRUE,
  raster=TRUE,
  raster_dpi=400, alpha=1, restrict_range=TRUE
)

pdf("figures/RGL_featureplot_MEs.pdf",height=10, width=15)
wrap_plots(plot_list, ncol=4)
dev.off()


################################################################################
# Network visualization
################################################################################

seurat_obj <- RunModuleUMAP(
  seurat_obj,
  n_hubs =5,
  n_neighbors=10,
  min_dist=0.5,
  spread=1,
  supervised=TRUE,
  target_weight=0.3
)

# get the hub gene UMAP table from the seurat object
umap_df <- GetModuleUMAP(
  seurat_obj
)

# plot with ggplot
p <- ggplot(umap_df, aes(x=UMAP1, y=UMAP2)) +
  geom_point(
   color=umap_df$color,
   size=umap_df$kME*2
  ) +
  umap_theme()

pdf(paste0(fig_dir, 'RGL_hubgene_umap_ggplot.pdf'), width=5, height=5)
p
dev.off()



library(reshape2)
library(igraph)
pdf(paste0(fig_dir, 'test_hubgene_umap_igraph2.pdf'), width=10, height=10)
ModuleUMAPPlot(
  seurat_obj,
  edge.alpha=0.5,
  sample_edges=TRUE,
  keep_grey_edges=FALSE,
  edge_prop=0.075, # taking the top 20% strongest edges in each module
  #label_genes = label_genes,
  label_hubs=3 # how many hub genes to plot per module?
)
dev.off()




  # plot with ggplot
  plot_df <- umap_df

  # compute coordinates for cluster labels
  centroid_df <- data.frame()
  for(cur_cluster in unique(plot_df[['module']])){
    cur_meta <- plot_df[plot_df[['module']] == cur_cluster,]
    df <- data.frame(
      cluster = cur_cluster,
      UMAP1 = mean(cur_meta$UMAP1),
      UMAP2 = mean(cur_meta$UMAP2)
    )
    centroid_df <- rbind(centroid_df, df)
  }


  # plot with ggplot
  p <- ggplot(umap_df, aes(x=UMAP1, y=UMAP2)) +
    ggrastr::rasterise(geom_point(
     color=umap_df$color,
     size=umap_df$kME*2
   ), dpi=500, scale=0.5) +
    umap_theme() +
    theme(
      plot.margin = margin(0,0,0,0),
      plot.title = element_text(hjust=0.5)
    ) +
    ggrepel::geom_text_repel(data = centroid_df, label=centroid_df$cluster, color='black', max.overlaps=Inf, size=3)


pdf(paste0(fig_dir, 'RGL_hubgene_umap_ggplot.pdf'), width=10, height=10)
p
dev.off()



# individual module networks
ModuleNetworkPlot(
  seurat_obj,
  mods = "all",
  #label_center=TRUE,
  outdir = paste0(fig_dir, 'RGL_hubNetworks/')
)




```


Save the results
```{r eval=FALSE}

saveRDS(seurat_obj, paste0(data_dir, 'sciso_RGL_hdWGCNA.rds'))

```

Run enrichr

```{r eval=FALSE}

library(enrichR)

dbs <-c('GO_Biological_Process_2021','GO_Cellular_Component_2021','GO_Molecular_Function_2021', 'WikiPathway_2021_Mouse', 'KEGG_2021_Mouse')

modules <- GetModules(seurat_obj); modules_full <- modules
modules <- dplyr::rename(modules, c(isoform_name = gene_name))
modules$gene_name <- sub("-[^-]+$", "", modules$isoform_name)
seurat_obj <- SetModules(seurat_obj, modules)

# compute GO terms:
seurat_obj <- RunEnrichr(seurat_obj, dbs=dbs)
seurat_obj <- SetModules(seurat_obj, modules_full)


enrichr_df <- GetEnrichrTable(seurat_obj) %>% subset(P.value < 0.05)
write.table(enrichr_df, quote=FALSE, sep='\t', row.names=FALSE, file=paste0(data_dir, 'RGL_enrichr.tsv'))

################################################################################
# Plot selected GO terms
################################################################################

modules <- GetModules(seurat_obj)
color_df <- modules %>% subset(module!='grey') %>%
  select(c(module, color)) %>% distinct %>%
  rename(c(group=module, colour=color))
mods <- levels(modules$module)
mods <- mods[mods!='grey']

# helper function to wrap text
wrapText <- function(x, len) {
    sapply(x, function(y) paste(strwrap(y, len), collapse = "\n"), USE.NAMES = FALSE)
}

combined_output <- GetEnrichrTable(seurat_obj)
selected_terms <- read.table('data/RGL_enrichr_selected.txt', sep='\t', header=1)

# subset selected terms
selected_terms <- subset(combined_output, Term %in% selected_terms$Term & P.value < 0.05)

selected_terms$group <- factor(
  as.character(selected_terms$module),
  levels = mods
)

# set max pval
quantile(-log(selected_terms$P.value), 0.95)
max_p <- 10

selected_terms$logp <- -log(selected_terms$P.value)
selected_terms$logp <- ifelse(selected_terms$logp > max_p, max_p, selected_terms$logp)

# remove GO Term ID
selected_terms$Term <- str_replace(selected_terms$Term, " \\s*\\([^\\)]+\\)", "")

selected_terms <- selected_terms %>%
  arrange(group)


selected_terms$wrap <- wrapText(selected_terms$Term, 35)

selected_terms$Term <- factor(
  as.character(selected_terms$Term),
  levels = rev(unique(as.character(selected_terms$Term)))
)

# GO Term dot plot
p <- selected_terms %>%
  ggplot(aes(x = group, y = Term, color =logp, size=log(Combined.Score))) +
  geom_point() +
  scale_color_stepsn(colors=rev(magma(256))) +
  RotatedAxis() + xlab('') + ylab('') +
  theme(
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    panel.border = element_rect(size=1, color='black', fill=NA),
    axis.line.x = element_blank(),
    axis.line.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    plot.margin = margin(0,0,0,0),
    panel.grid = element_line(size=0.25, color='lightgrey')
  )


# make the colorbar as its own heatmap
color_df$var <- 1
colorbar <- color_df %>%
  ggplot(aes(x=group, y=var, fill=group)) +
  geom_tile() +
  scale_fill_manual(values=color_df$colour) +
  coord_equal() +
  NoLegend() + RotatedAxis() +
  theme(
    plot.title=element_blank(),
    axis.line=element_blank(),
    axis.ticks.y =element_blank(),
    axis.text.y = element_blank(),
    axis.title = element_blank(),
    plot.margin=margin(0,0,0,0),
  )

pdf(paste0(fig_dir, 'selected_GO_terms.pdf'), width=9, height=8)
p / colorbar #+ plot_layout(heights=c(20,1))
dev.off()

```

Identify genes with isoforms in more than one module

```{r eval=FALSE}

modules <- GetModules(seurat_obj)
color_df <- modules %>% subset(module!='grey') %>%
  select(c(module, color)) %>% distinct %>%
  rename(c(group=module, colour=color))
cp <- color_df$colour; names(cp) <- color_df$group

modules <- dplyr::rename(modules, c(isoform_name = gene_name))
modules$gene_name <- sub("-[^-]+$", "", modules$isoform_name)
modules <- subset(modules, module != 'grey')

write.csv(modules, row.names=FALSE, quote=FALSE, file=paste0(data_dir, 'RGL_isoseq_modules.csv'))

mod_iso <- table(modules$gene_name, modules$module)
mod_iso[mod_iso > 1] <- 1

multimod_genes <- names(which(Matrix::rowSums(mod_iso) > 1))

multimod_df <- subset(modules, gene_name %in% multimod_genes)

subset(multimod_df, gene_name == 'Gfap')

# write it to a file:
multimod_df %>% dplyr::select(c(isoform_name, gene_name, module, color)) %>%
  write.csv(file=paste0(data_dir, 'multimod_genes.csv'), quote=FALSE, row.names=FALSE)


################################################################################
# how many genes have 1 isoform, all isoforms in the same module, isoforms split
# across modules?
################################################################################

# total number of genes:
n_genes <- modules$gene_name %>% unique %>% length

# number of genes with one isoform:
one_iso <- names(which(table(modules$gene_name) == 1))

n_one <- length(one_iso)
n_multi_split <- length(multimod_genes)
n_multi_same <- n_genes - n_one - n_multi_split

data <- data.frame(
  category = c('one isoform', 'multi iso, same module', 'mutli iso, diff modules'),
  count = c(n_one, n_multi_same, n_multi_split)
)

# Compute percentages
data$fraction <- data$count / sum(data$count)

# Compute the cumulative percentages (top of each rectangle)
data$ymax <- cumsum(data$fraction)

# Compute the bottom of each rectangle
data$ymin <- c(0, head(data$ymax, n=-1))

# Compute label position
data$labelPosition <- (data$ymax + data$ymin) / 2

# percentage:
data$percent <- paste0(signif(data$fraction * 100, 3), "%")

# Compute a good label
data$label <- paste0(data$category, "\n", data$percent)

# Make the plot
p <- ggplot(data, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=category)) +
  geom_rect() +
  geom_text( x=2, aes(y=labelPosition, label=label), size=5, color='black') +
  scale_fill_brewer(palette=4) +
  coord_polar(theta="y") +
  xlim(c(0, 4)) +
  theme_void() +
  theme(legend.position = "none")


pdf(paste0(fig_dir, 'RGL_donut_multimod.pdf'), width=5, height=5)
p
dev.off()


################################################################################
# network vis with selected genes
################################################################################

hubs <- GetHubGenes(seurat_obj, 25)

# which genes are hub genes in more than one module?
hubs <- dplyr::rename(hubs, c(isoform_name = gene_name))
hubs$gene_name <- sub("-[^-]+$", "", hubs$isoform_name)
hubs <- hubs %>% dplyr::select(c(gene_name, module)) %>% distinct()
multimod_hubs <- names(which(table(hubs$gene_name) > 1))
multimod_hubs

plot_genes <- c(
  'Gfap-0', 'Gfap-1', 'Gfap-2', 'Gfap-3', 'Gfap-5',
  'Cd9-0', 'Cd9-1',
  'Cd81-0', 'Cd81-1',
  'Fabp5-0', 'Fabp5-1',
  'H3f3b-0', 'H3f3b-1'
)

library(igraph)
pdf(paste0(fig_dir, 'RGL_coex_umap_multimod_igraph2.pdf'), width=7, height=7)
ModuleUMAPPlot(
  seurat_obj,
  edge.alpha=0.5,
  sample_edges=TRUE,
  keep_grey_edges=FALSE,
  edge_prop=0.075, # taking the top 20% strongest edges in each module
#   label_genes = umap_df$isoform_name[umap_df$gene_name %in% plot_genes],
  label_genes = plot_genes,
  label_hubs=2 # how many hub genes to plot per module?
)
dev.off()



```

Run UMAP on RGL alone

```{r eval=FALSE}

library(MetBrewer)
new_colors <- paste0(met.brewer("Lakota", n=4, type='discrete'))
color_df <- data.frame(
  group = unique(seurat_obj$annotation),
  colour = new_colors
)
cp <- color_df$colour; names(cp) <- color_df$group

DefaultAssay(seurat_obj) <- 'RNA'
seurat_obj <- RunUMAP(seurat_obj, dims=1:15, min.dist=0.2, n.neighbors=15L, spread=0.5)

p <- PlotEmbedding(
  seurat_obj,
  group.by = 'annotation',
  label=TRUE,
  raster=TRUE,
  raster_dpi = 600,
  point_size=1,
  raster_scale=0.75,
  plot_under=TRUE,
  plot_theme = umap_theme(),
  color_df = color_df
)


pdf(paste0(fig_dir, 'RGL_umap.pdf'), width=5, height=5)
p
dev.off()

# umap of full dataset
p <- PlotEmbedding(
  seurat_full,
  group.by = 'annotation',
  label=TRUE,
  raster=TRUE,
  raster_dpi = 600,
  point_size=1,
  raster_scale=0.5,
  plot_under=TRUE,
  plot_theme = umap_theme(),
  color_df = color_df,
  selected = c('RGL', 'ASC', 'Ependymal', 'NIPCs')
)

pdf(paste0(fig_dir, 'RGL_umap_full.pdf'), width=5, height=5)
p
dev.off()


```

Pseudotime trajectory analysis with monocle3

```{r eval=FALSE}

library(monocle3)
library(SeuratWrappers)

# process the CDS
rgl_cds <- as.cell_data_set(seurat_obj)
rgl_cds <- cluster_cells(cds=rgl_cds, reduction_method='UMAP')


p <- plot_cells(
  cds = rgl_cds,
  color_cells_by = "partition",
  show_trajectory_graph = FALSE
)

pdf(paste0(fig_dir, 'rgl_umap_monocle.pdf'),  width=7, height=5, useDingbats=FALSE)
p
dev.off()

# learn graph for pseudotime
rgl_cds <- learn_graph(
  rgl_cds,
  use_partition=FALSE,
  learn_graph_control=list(prune_graph=FALSE)
)

p <- plot_cells(
  cds = rgl_cds,
  color_cells_by = "annotation",
  show_trajectory_graph = TRUE,
  label_principal_points = TRUE
) + umap_theme()

pdf(paste0(fig_dir, 'rgl_principal_points.pdf'),  width=5, height=5, useDingbats=FALSE)
p
dev.off()

# get principal node & order cells based on where the radial glia are as the start
principal_node <- 'Y_7'
rgl_cds <- order_cells(rgl_cds,root_pr_nodes = principal_node)

# add pseudotime to seurat object:
seurat_obj$pseudotime <- pseudotime(rgl_cds)
seurat_obj$asc_pseudotime <- ifelse(seurat_obj$annotation %in% c("RGL", "ASC"), seurat_obj$pseudotime, NA)
seurat_obj$epd_pseudotime <- ifelse(seurat_obj$annotation %in% c("RGL", "Ependymal"), seurat_obj$pseudotime, NA)
seurat_obj$nipc_pseudotime <- ifelse(seurat_obj$annotation %in% c("RGL", "NIPCs"), seurat_obj$pseudotime, NA)

p <- plot_cells(
  cds = rgl_cds,
  color_cells_by = "pseudotime",
  show_trajectory_graph = TRUE,
  label_roots = TRUE,
  label_leaves = FALSE,
  label_branch_points = FALSE,
) + umap_theme()

pdf(paste0(fig_dir, 'umap_pseudotime.pdf'),  width=5, height=4, useDingbats=FALSE)
p
dev.off()

p1 <- FeaturePlot(seurat_obj, feature='asc_pseudotime', raster=FALSE) + umap_theme() + scale_color_gradientn(colors=plasma(256)) + coord_equal()
pdf(paste0(fig_dir, 'test_umap_pseudotime.pdf'),  width=4, height=4, useDingbats=FALSE)
p1
dev.off()


saveRDS(seurat_obj, paste0(data_dir, 'sniso_RGL_hdWGCNA.rds'))

```

Plot pseudotime trajectories

```{r eval=FALSE}

library(MetBrewer)
new_colors <- paste0(met.brewer("Lakota", n=4, type='discrete'))
color_df <- data.frame(
  group = unique(seurat_obj$annotation),
  colour = new_colors
)
cp <- color_df$colour; names(cp) <- color_df$group

seurat_obj$UMAP1 <- seurat_obj@reductions$umap@cell.embeddings[,1]
seurat_obj$UMAP2 <- seurat_obj@reductions$umap@cell.embeddings[,2]

# Ependymal
p1 <- seurat_obj@meta.data %>%
  ggplot(aes(x=UMAP1, y=UMAP2, color=epd_pseudotime)) +
  ggrastr::rasterise(geom_point(size=1), dpi=500, scale=0.75) +
  coord_equal() +
  scale_color_gradientn(colors=plasma(256), na.value='grey') +
  umap_theme()

# ASC
p2 <- seurat_obj@meta.data %>%
  ggplot(aes(x=UMAP1, y=UMAP2, color=asc_pseudotime)) +
  ggrastr::rasterise(geom_point(size=1), dpi=500, scale=0.75) +
  coord_equal() +
  scale_color_gradientn(colors=viridis(256), na.value='grey') +
  umap_theme()

# NPCs
p3 <- seurat_obj@meta.data %>%
  ggplot(aes(x=UMAP1, y=UMAP2, color=nipc_pseudotime)) +
  ggrastr::rasterise(geom_point(size=1), dpi=500, scale=0.75) +
  coord_equal() +
  scale_color_gradientn(colors=inferno(256), na.value='grey') +
  umap_theme()

pdf(paste0(fig_dir, 'RGL_umap_pseudotime.pdf'),  width=8, height=8, useDingbats=FALSE)
p1 + p2 + p3 + plot_layout(ncol=2, guides='collect')
dev.off()

```

Module Trajectories

```{r eval=FALSE}

# cut pseudotime into bins using galilean:
seurat_obj <- BinPseudotime(seurat_obj, pseudotime_col='nipc_pseudotime', n_bins=c(10, 20))
seurat_obj <- BinPseudotime(seurat_obj, pseudotime_col='asc_pseudotime', n_bins=c(10, 20))
seurat_obj <- BinPseudotime(seurat_obj, pseudotime_col='epd_pseudotime', n_bins=c(10, 20))

MEs <- GetMEs(seurat_obj)
modules <- GetModules(seurat_obj)
mods <- levels(modules$module)
mods <- mods[mods!='grey']
module_colors <- modules %>% dplyr::select(c(module, color)) %>% distinct()
rownames(module_colors) <- module_colors$module
mod_colors <- module_colors[mods, 'color']

seurat_obj@meta.data <- cbind(seurat_obj@meta.data, MEs)


################################################################
# EPD pseudotime
################################################################

# compute the average ME score in each bin:
features <- mods

avg_scores <- seurat_obj@meta.data %>%
  group_by(epd_pseudotime_bins_20) %>%
  select(all_of(features)) %>%
  summarise_all(mean)

colnames(avg_scores)[1] <- 'bin'

plot_df <- reshape2::melt(avg_scores)
epd_df <- plot_df

p1 <- ggplot(plot_df, aes(x = as.numeric(bin), y = value, color=variable)) +
  geom_point() +
  geom_smooth() +
  scale_color_manual(values=mod_colors)+
  xlab('Pseudotime') +
  ylab('ME') +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  NoLegend()

pdf(paste0(fig_dir, 'EPD_pseudotime_MEs.pdf'), width=10, height=6)

p1 + facet_wrap(~variable, ncol=4, scales='free')

dev.off()


################################################################
# ASC pseudotime
################################################################

# compute the average ME score in each bin:
features <- mods

avg_scores <- seurat_obj@meta.data %>%
  group_by(asc_pseudotime_bins_20) %>%
  select(all_of(features)) %>%
  summarise_all(mean)

colnames(avg_scores)[1] <- 'bin'

plot_df <- reshape2::melt(avg_scores)
asc_df <- plot_df

p1 <- ggplot(plot_df, aes(x = as.numeric(bin), y = value, color=variable)) +
  geom_point() +
  geom_smooth() +
  scale_color_manual(values=mod_colors)+
  xlab('Pseudotime') +
  ylab('ME') +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  NoLegend()

pdf(paste0(fig_dir, 'ASC_pseudotime_MEs.pdf'), width=10, height=6)
p1 + facet_wrap(~variable, ncol=4, scales='free')
dev.off()

################################################################
# NPC pseudotime
################################################################

# compute the average ME score in each bin:
features <- mods

avg_scores <- seurat_obj@meta.data %>%
  group_by(nipc_pseudotime_bins_20) %>%
  select(all_of(features)) %>%
  summarise_all(mean)

colnames(avg_scores)[1] <- 'bin'

plot_df <- reshape2::melt(avg_scores)
nipc_df <- plot_df

p1 <- ggplot(plot_df, aes(x = as.numeric(bin), y = value, color=variable)) +
  geom_point() +
  geom_smooth() +
  scale_color_manual(values=mod_colors)+
  xlab('Pseudotime') +
  ylab('ME') +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  ) +
  NoLegend()

pdf(paste0(fig_dir, 'NIPCs_pseudotime_MEs.pdf'), width=10, height=6)
p1 + facet_wrap(~variable, ncol=4, scales='free')
dev.off()


################################################################
# Plot trajectories for all modules with all 3 groups:
################################################################

asc_df$bin <- as.numeric(asc_df$bin); asc_df$group <- 'ASC'
epd_df$bin <- as.numeric(epd_df$bin); epd_df$group <- 'Ependymal'
nipc_df$bin <- as.numeric(nipc_df$bin); nipc_df$group <- 'NIPCs'

plot_df <- do.call(rbind, list(asc_df, epd_df, nipc_df))

p1 <- plot_df %>%
  ggplot(aes(x = as.numeric(bin), y = value, color=group)) +
  geom_smooth(se=FALSE) +
  geom_hline(yintercept=0, linetype='dashed', color='grey') +
  scale_color_manual(values=cp[unique(plot_df$group)])+
  xlab('Pseudotime') +
  ylab('ME') +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank()
  )

pdf(paste0(fig_dir, 'RGL_pseudotime_MEs.pdf'), width=6, height=5)
p1 + facet_wrap(~variable, ncol=3, scales='free')
dev.off()



```

save data for Swan

```{r eval=FALSE}

X <- GetAssayData(seurat_full, slot='counts', assay='iso')
df <- as.data.frame(X)

zz <- gzfile("/pub/smorabit/hdWGCNA/isoseq/data/dataSharing_countsPerCell/isoformByCellMatrix/features.tsv.gz", 'rt')
feats <- read.table(zz)

rownames(df) <- feats$V1
bcs <- colnames(df)
df$transcript_id <- feats$V1
df <- df[,c('transcript_id', bcs)]

write.table(df, quote=FALSE, file=paste0(data_dir, 'sniso_counts.tsv'), sep = '\t', row.names=FALSE)
write.table(seurat_full@meta.data, quote=FALSE, file=paste0(data_dir, 'sniso_cell_meta.tsv'))

```

one-versus-all Differential module eigengene (DME) analysis

```{r eval=FALSE}

DMEs_all <- FindAllDMEs(
  seurat_obj,
  group.by = 'annotation',
)

write.csv(DMEs_all, row.names=FALSE, quote=FALSE, file=paste0(data_dir, 'RGL_DMEs.csv'))

```
